---
title: "Data processing"
author: "James Ashmore"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Setup

Set chunk options:

```{r}
knitr::opts_chunk$set(
  autodep    = TRUE,
  cache      = TRUE,
  cache.path = "cache/02-data-processing.Rmd/",
  dev        = "png",
  error      = FALSE,
  message    = FALSE,
  warning    = FALSE
)
```

Load required packages:

```{r}
library(batchelor)
library(scater)
library(scran)
library(EnsDb.Mmusculus.v79)
library(org.Mm.eg.db)
library(msigdbr)
library(pheatmap)
library(viridis)
library(ggthemes)
```

## Import data

```{r}
sce <- readRDS("output/01-data-loading.rds")
```

## Quality control

Quality control on the cells has already been performed by the authors, so we will just use their filtering :

```{r}
sce <- sce[, sce$use]
```

## Normalization

```{r}
sce <- logNormCounts(sce)
```

## Blacklist

Create a blacklist of genes to ignore for feature selection:

```{r}
rowData(sce)$blacklist <- FALSE
```

Append Xist gene to blacklist:

```{r}
idx <- which(rowData(sce)$gene_id == "ENSMUSG00000086503")

rowData(sce)$blacklist[idx] <- TRUE
```

Append Y chromosome genes to blacklist:

```{r}
nan <- is.na(rowData(sce)$gene_id)

loc <- mapIds(EnsDb.Mmusculus.v79, keys = rowData(sce)$gene_id[!nan], keytype = "GENEID", column = "SEQNAME")

loc <- loc[loc == "Y"]

idx <- which(rowData(sce)$gene_id %in% names(loc))

rowData(sce)$blacklist[idx] <- TRUE
```

Append MT chromosome genes to blacklist:

```{r}
nan <- is.na(rowData(sce)$gene_id)

loc <- mapIds(EnsDb.Mmusculus.v79, keys = rowData(sce)$gene_id[!nan], keytype = "GENEID", column = "SEQNAME")

loc <- loc[loc == "MT"]

idx <- which(rowData(sce)$gene_id %in% names(loc))

rowData(sce)$blacklist[idx] <- TRUE
```

Append cell-cycle genes to blacklist:

```{r}
org <- select(org.Mm.eg.db, keys = "GO:0007049", keytype = "GOALL", column = "ENSEMBL")

idx <- which(rowData(sce)$gene_id %in% org$ENSEMBL)

rowData(sce)$blacklist[idx] <- TRUE
```

Append ribosomal genes to blacklist:

```{r}
sig <- msigdbr(species = "Mus musculus", category = "C2")

sig <- subset(sig, gs_name == "KEGG_RIBOSOME")

org <- select(org.Mm.eg.db, keys = sig$gene_symbol, keytype = "SYMBOL", column = "ENSEMBL")

idx <- which(rowData(sce)$gene_id %in% org$ENSEMBL)

rowData(sce)$blacklist[idx] <- TRUE
```

## Variance modelling

```{r}
dec <- modelGeneVarByPoisson(sce, block = sce$sample)

hvg <- dec$bio > 0

nan <- rowData(sce)$blacklist

hvg <- hvg & !nan
```

## Merging

```{r}
set.seed(1701)

mnn <- correctExperiments(sce, 
    batch = sce$sample, 
    subset.row = hvg,
    correct.all = TRUE,
    PARAM = FastMnnParam(auto.merge = TRUE)
)

mnn$batch <- factor(mnn$batch)
```

## Clustering

```{r}
snn <- buildSNNGraph(mnn, use.dimred = "corrected")

com <- igraph::cluster_louvain(snn)

mnn$cluster <- factor(com$membership)
```

## Dimensionality reduction

```{r}
mnn <- runTSNE(mnn, dimred = "corrected", external_neighbors = TRUE)

mnn <- runUMAP(mnn, dimred = "corrected", external_neighbors = TRUE)
```

## Cell cycle

```{r}
set.seed(1701)

mcm <- readRDS(system.file("exdata", "mouse_cycle_markers.rds", package = "scran"))

fit <- cyclone(mnn, mcm, gene.names = rowData(mnn)$gene_id)

mnn$phase <- factor(fit$phases)
```

## Data exploration

Create plotting data:

```{r}
dim <- reducedDim(mnn, "UMAP")

dim <- as.data.frame(dim)

colnames(dim) <- c("x", "y")

dat <- colData(mnn)

dat <- cbind(dim, dat)
```

Plot cells coloured by phase:

```{r}
phase_palette <- tableau_color_pal(palette = "Traffic")

phase_nlevels <- nlevels(mnn$phase)

phase_palette <- phase_palette(phase_nlevels)

names(phase_palette) <- levels(mnn$phase)

ggplot(dat, aes(x, y, colour = phase)) + 
  geom_point(data = dim, aes(x, y), colour = "gainsboro") + 
  geom_point() + 
  scale_colour_manual(name = "Phase", values = phase_palette) + 
  facet_wrap(~ batch) + 
  theme_void() + 
  theme(aspect.ratio = 1)
```

Plot cells coloured by batch:

```{r}
mnn$batch <- factor(mnn$batch)

batch_palette <- tableau_color_pal(palette = "Tableau 10")

batch_nlevels <- nlevels(mnn$batch)

batch_palette <- batch_palette(batch_nlevels)

names(batch_palette) <- levels(mnn$batch)

ggplot(dat, aes(x, y, colour = batch)) + 
  geom_point(data = dim, aes(x, y), colour = "gainsboro") + 
  geom_point() + 
  scale_colour_manual(name = "Batch", values = batch_palette) + 
  facet_wrap(~ batch) + 
  theme_void() + 
  theme(aspect.ratio = 1)
```

Plot cells coloured by cluster:

```{r}
cluster_palette <- tableau_color_pal(palette = "Tableau 20")

cluster_nlevels <- nlevels(mnn$cluster)

cluster_palette <- cluster_palette(cluster_nlevels)

names(cluster_palette) <- levels(mnn$cluster)

ggplot(dat, aes(x, y, colour = cluster)) + 
  geom_point(data = dim, aes(x, y), colour = "gainsboro") + 
  geom_point() + 
  scale_color_manual(name = "Cluster", values = cluster_palette) +  
  facet_wrap(~ batch) + 
  theme_void() + 
  theme(aspect.ratio = 1)
```

## Output

Save SCE object to disk:

```{r}
saveRDS(mnn, file = "output/02-data-processing.rds")
```

Print session information

```{r}
sessionInfo()
```
