---
title: "Interactive data exploration"
author: "James Ashmore"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Setup

Set chunk options:

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

Load CRAN packages:

```{r}
library(rsconnect)
```

Include Bioconductor package repositories:

```{r}
options(repos = BiocManager::repositories())
```

Automatically answer yes to any user prompt:

```{r}
options(needs.promptUser = FALSE)
```

Delete existing shiny directory:

```{r}
unlink("shiny", recursive = TRUE)
```

Create shiny directory for interactive data exploration:

```{r}
dir.create("shiny", showWarnings = FALSE, mode = "0755")
```

## Applications {.tabset}

### Processed data

Create iSEE directory for processed data:

```{r}
dir.create("shiny/processed", showWarnings = FALSE, mode = "0755")
```

Copy processed data to iSEE directory:

```{r}
file.copy("output/02-data-processing.rds", "shiny/processed/data.rds", overwrite = TRUE)
```

Write code to iSEE application file:

```{cat, engine.opts = list(file = "shiny/processed/app.R"), class.source = "r"}
# Load required packages
library(iSEE)
library(SingleCellExperiment)
library(BiocSingular)

# Read experiment object
sce <- readRDS("data.rds")

# Rename brachyury gene name
ind <- which(rowData(sce)$gene_id == "ENSMUSG00000062327")
rowData(sce)$gene_name[ind] <- "Bra"

# Make feature names unique
rownames(sce) <- scuttle::uniquifyFeatureNames(rowData(sce)$gene_id, rowData(sce)$gene_name)

# Identify NMP cells
ids <- list(
  "Sox2" = counts(sce)["Sox2", ] > 0,
  "Bra" = counts(sce)["Bra", ] > 0
)
sce$NMP <- ids$Sox2 & ids$Bra

# Identify preNeuro cells
ids <- list(
  "Sox2" = counts(sce)["Sox2", ] > 0,
  "Nkx1-2" = counts(sce)["Nkx1-2", ] > 0,
  "Bra" = counts(sce)["Bra", ] > 0
)
sce$preNeuro <- (ids$Sox2 & ids$`Nkx1-2`) & !ids$Bra

# Identify preMeso cells
ids <- list(
  "Bra" = counts(sce)["Bra", ] > 0,
  "Sox2" = counts(sce)["Sox2", ] > 0,
  "Mixl1" = counts(sce)["Mixl1", ] > 0,
  "Bmp4" = counts(sce)["Bmp4", ] > 0
)
sce$preMeso <- ids$Bra & !(ids$Sox2 | ids$Mixl1 | ids$Bmp4)

# Identify LPM cells
ids <- list(
  "Hand1" = counts(sce)["Hand1", ] > 0,
  "Foxf1" = counts(sce)["Foxf1", ] > 0
)
sce$LPM <- ids$Hand1 & ids$Foxf1

# Set iSEE options
iSEEOptions$set(downsample = TRUE)
iSEEOptions$set(color.maxlevels = 37)

# Create iSEE object
iSEE(sce, appTitle = "CLE_NMP")
```

Deploy the iSEE application to the shiny server:

```{r}
deployApp(appDir = "shiny/processed", appName = "CLE_NMP", appTitle = "CLE_NMP", launch.browser = FALSE, forceUpdate = TRUE)
```

Allocate more memory to the iSEE application:

```{r}
configureApp(appName = "CLE_NMP", appDir = "shiny/processed", size = "xxlarge")
```

## Summary

### Session

Print session information:

```{r}
sessionInfo()
```
